---
title: "Does household income affect turnout?"


---

## Introduction

This was actually a something I was thinking of while doing the recent Johor election prediction (but couldn't realize due to work and time constraints). 

All election data comes from [ElectionData.MY](https://electiondata.my){target="_blank"}, the public website of the Malaysian Election Corpus (MECo;
[Scientific Data, 2025](https://www.nature.com/articles/s41597-025-06502-7){target="_blank"}). The site dedicates its data to the public domain (CC0).

All code used can be found in this [GitHub repo](https://github.com/mary-emm/mm-portfolio/tree/main/projects/johor-turnout){target="_blank"}. A short write up can also be [here](#){target="_blank"}.

## Why do I care about this question?

When I joined INVOKE, I knew very little about how politics worked, and I had the impression that most people my age didn't care about it. That was humbling. I still however, **keep my distance from politics itself**, but I don't think the goal is to immerse yourself in it. It's to understand it well enough to make decisions that are reasoned, not just emotional. I treat this exploration as factual answers to my questions that I otherwise won't have known unless I did this.

## Load Johor's SE-15 and compute turnout
```{r}
#| message: false
library(readxl)
library(dplyr)
library(ggplot2)



se_15 <- read_excel("data/raw/jhr_se15_dun_turnout.xlsx") %>%
  mutate(turnout = ballots_issued / voters_total)


glimpse(se_15)
summary(se_15$turnout)
```


## Load the income & income inequality data

Since both files covers between 2019-2024, I'll filter out only Johor & 2022.
```{r}
income_raw <- read.csv("data/raw/hh_income_dun.csv")

glimpse(income_raw)
income_raw %>% distinct(date)


# Since we're looking at Johor SE-15 (2022), I'll filter out Johor and the income data from 2022
income <- income_raw %>% filter(state == "Johor", substr(date,1,4) == "2022")

glimpse(income)

```

```{r}
ineq_raw <- read.csv("data/raw/hh_inequality_dun.csv")

glimpse(ineq_raw)
ineq_raw %>% distinct(date)


# Since we're looking at Johor SE-15 (2022), I'll filter out Johor and the income data from 2022
ineq <- ineq_raw %>% filter(state == "Johor", substr(date,1,4) == "2022")

glimpse(ineq)

```

## Checking if the DUN name in the household income matches the one from ElectionData

```{r}
setdiff(se_15$seat, income$dun)
setdiff(se_15$seat, ineq$dun)
```

Matches so it makes things easier.

## Joining the election data with the income + inequality data.
```{r}

jhr <- se_15 %>% left_join(income, by = c("seat" = "dun")) %>% left_join(ineq, by = c("seat" = "dun")) %>% select(election,seat,voters_total, ballots_issued,ethnic_malay,ethnic_chinese,ethnic_indian,turnout, income_median,income_mean,gini)

glimpse(jhr)

colSums(is.na(jhr))
```


```{r}

ggplot(jhr, aes(x=income_median,y= turnout)) + geom_point() + geom_smooth(method= "lm") + labs(
  title = "Median income vs Turnout (Johor SE-15)",
    x = "Median income",
    y = "Avg. Turnout"
)
```


```{r}

ggplot(jhr, aes(x=gini,y= turnout)) + geom_point() + geom_smooth(method= "lm") + labs(
  title = "Income inequality (Gini) vs Turnout (Johor SE-15)",
    x = "Income inequality",
    y = "Avg. Turnout"
)


cor(jhr$turnout,jhr$gini)
cor(jhr$turnout, jhr$income_median)
```

Median income has clearer negative correlation, 

```{r}

m1 <- lm(turnout ~ I(income_median / 1000), data = jhr)

summary(m1)
```


