---
title: "Was it worth the flip?"
categories: [Personal, 2026, R]
---

## Background

Being very late to the party when it comes to card games, I have since found a fascination towards them. Specifically, a card game called `Flip 7`, where the goal of the game is to flip 7 **distinct** cards, without busting.

As simple as it sounds, it's a probabilistic game. Some would say luck, which it is still, since it's kinda hard to keep track of how many cards are left. That's what makes it interesting: the odds change with every card that leaves the deck.

## How the game works

Everyone flips cards from a shared deck, one at a time. The deck has number cards from 0 to 12, and there are as many copies of each number as its value (one 1, two 2s, and so on). Flip a number you already hold and you bust, scoring nothing that round. Stop before that happens and you bank the total of your hand. Flip seven *different* numbers and you earn a bonus, and the first player to 200 points wins. There are also action cards, such as **Freeze**, **Flip Three** and **Second Chance**.

## This project

To keep things simple, I only use the number cards 0 to 12 (79 cards) and leave out the bonus cards. The question I want to answer is: **when to stop flipping?**

Please find the write-up to this [here](test){target="_blank"}


## Code


### Create the deck
```{r}
#| message: false

library(tidyverse)

# create card deck containing 0-5 numbers (for 0 and 1 card, theres 1 each)

deck <- c(0, 1, rep(2,2), rep(3,3), rep(4,4), rep(5,5),
          rep(6,6), rep(7,7), rep(8,8), rep(9,9),
          rep(10,10), rep(11,11), rep(12,12))

deck

# should be 79
length(deck)
```


### Simulate the game

```{r}
play_round <- function(stop_at){
  
  # randomly shuffle deck
  d <- sample(deck)
  
  # initializing hand and total
  hand <- c()
  total <- 0
  
  for (card in d){
    # the total reaches the limit, then stop
    if (total >= stop_at) return(total)
    
    # if pull the same card, then bust
    if (card %in% hand) return(0)
    hand <- c(hand, card)
    total <- total + card
    
    # whoever gets 7 distinct card gets an additional 15 points
    if (length(hand)==7) return(total +15)
  }
  total
  

  }
```

### Test simulation

```{r}
# test 
replicate(10, play_round(20))
```


```{r}

set.seed(47)
n <- 5000

results <- map_dfr(1:63, function(s) {
  scores <- replicate(n, play_round(s))
  tibble(stop_at = s,
         avg_points = mean(scores),
         se = sd(scores) / sqrt(n))
})

head(results)
```


```{r}
#| message: false
#| fig-width: 7
#| fig-height: 4

best <- results %>% slice_max(avg_points, n = 1, with_ties = FALSE)

results %>%
  ggplot(aes(stop_at, avg_points)) +
  geom_ribbon(aes(ymin = avg_points - 2 * se, ymax = avg_points + 2 * se),
              alpha = 0.2) +
  geom_line() +
  geom_point(size = 1) +
  geom_point(data = best, colour = "red", size = 3) +
  geom_vline(xintercept = best$stop_at, linetype = "dashed", colour = "red") +
  geom_label(
    data = best,
    aes(label = paste0("Best: stop at ", stop_at, "\n", round(avg_points, 2), " pts")),
    vjust = -0.3, colour = "red", size = 3.5
  ) +
  expand_limits(y = max(results$avg_points) * 1.2) +
  labs(x = "Stop once total reaches", y = "Average points per round") +
  theme_minimal(base_size = 11)
```


```{r}
# examining which card is around the same as stopping when point exceeds 1
results %>% filter(stop_at %in% c(1, 40:50)) %>% select(stop_at, avg_points, se)
```
